Neural Network Language Models for Automatic Speech Recognition

### 2018-12-13 paper/paper.tex
paper: Throw-away line about why we care about the data
review of paper and fixing the bibliography

Prior to the introduction of RNNs, most language models used n-grams
in Hidden Markov Model chains, where it was assumed that the $k$th
word depended only on the prior $n$ words. One key problem
with this model is that the context window is of a fixed size, leading
to a horizon problem. The horizon problem occurs because a if there were a word
in the word sequence which is extremely predictive of the $(n + 1)$th
word away from the current word, that word would simply not be taken
into account in that model. Worse still, the fixed size context window
means that unnecessary weight is put on words that happen to be in the context
window that may actually not be all that predictive were the context window
infinite.

In contrast, RNN's seek to solve this problem by introducing a Neural Network
architecture with recurrent connections, meaning that as the RNN makes predictions
over a sequence observations, it takes into account a hidden layer produced
by the previous observation (which in turn was a product of all the observations
before that). In effect, the RNN architecture encodes the contextual history
into the hidden layer vector allowing for an encoded approximation of infinite
contextual history.

@@

@@

The first dataset we use is the WikiText dataset. This is a smaller dataset than
Gigaword and WSJ'92, but allowed us to
iterate faster and run more experiments.

@@

$$ WER = \frac{S + D + I}{N} $$, where $N$ is the number of
words and $S, D, I$ are substitutions, deletions, insertions.
This is similar to the Levensteihn distance between documents.

However, such a metric doesn't take int account the fact that
a recognized sequence may differ in length to a reference
sequence.

Improving on this situation, ROUGE measures based on the
longest matching subsequences, overlapping pairs and n-gram
co-occurrences.

\subsection{METEOR}
\label{sec:meteor}
Another metric trying to improve BLEU is
METEOR. The novel thing about this metric is the use of Recall as well as
Precision to compute the score. It computes the harmonic mean of Recall and
Precision over uni-grams, the weight for recall being higher than Precision,
meaning that false negatives (having a low probability for the true word trying
to predict) will have a big impact on the metric (the score will be lower as the
amount of false negatives increase).

More formally, Precision over an uni-gram is computed as: $$P={\frac
{m}{w_{{t}}}}$$ Where m is the number of uni-grams in the candidate translation
that are also found in the reference translation, and $w_t$ is the number of
uni-grams in the candidate translation.

@@

Unfortunately, Elman Network RNNs are not able to retain all the optimal
contextual in practice. The reasons for this are given by Bengio\cite{Bengio94},
but in summary the issue is that as we backpropagate through time the gradients
from states "further back" in the history start to get vanishingly small, meaning
that they can be easily disrupted by noise or newer information. This means
that the hidden layer doesn't retain long-term information since it is easily
disrupted by new information.

@@

The LSTM cell uses four "gates" to manage the state of the memory. Each of these
gates comprises a set of weights and activations, descibed below \cite{colah2015}.

@@

The last gate is an "output" gate which decides which parts of the hidden state are
taken on account to compute the output. The hidden state is first activated by sigmoid to determin
the mask ($o_t$), then the hidden layer is activated by tanh to push the values between $[-1, 1]$, then
finally multipled by the mask $o_t$ in order to determine the output of the network.

The end result with such a variation is that the hidden state is "protected" from perturbation
by virtue of being combined through a single linear layer and activation layer with the input
state on every timestep. Instead, the hidden state is more "explicitly" modified when the
optimized input and forget gates explicitly indicate that information should be added or removed
to the state \cite{hochreiter97}.

@@

Note that the hidden state accuracy benefits here come with a computational compleixty cost - every
LSTM cell has three internal linear layers and five activations, which we pay for both in the forward
pass and in backpropagation through time.

@@

Traditionally, the Language Models were feeded each word as a One-Hot encoded vector as large as the
vocabulary is. Since vocabularies can be really large (20k to 800k for languages
with many compounded words such as Finnish), this presents a computational
problem, large matrices multiplication is expensive. Word Embeddings provide an
alternative procedure in which the words to be fed to the network are encoded
in an $n$ dimensional space. The embedding layer can learn a word encoding such
that words related to each other semanticaly, get similar encoded vectors after
the transformation. Even some basic arithmetic is possible by adding and subtracting the euclidean distance of vectors. For example, if you
subtract (the vector for "queen" - the vector from "woman") from the vector for "king" you get the vector
for "man".

@@

Word embeddings also have the distinct advantage that they can reduce both training
and inference complexity because fewer multiplications need to be done at
the RNN, LSTM or GRU cell phase both at the inference and training time. The
comptutation of the embedding layer can be heavily optimized because the relevant
input will only have a non-zero value in a single vector component.

@@

The overall theory behind shortlists is that any optimal model operates
on a pareto principle, 20% of the words which are the most frequent will
be predicted roughly 80% of the time. Since the computation of the output
encodings of height $|V|$ is a fixed cost on every training and inference
cycle, the user of the model is paying for the capacity to predict words which
will seldomly be predicted. A good shortlist should be able to substantially
reduce the computational complexity of the model whilst still retaining a
relatively high degree of accuracy.

@@

The Milkolov paper flags that performance on the validation could not be
improved by using "standard regularization procedures" (we assume that
the paper must have been referring to L1 or L2 regualrization based on the
techniques in existence at the time) \cite{Milkolov10}. We propose that a
popular regularization technique known as "Dropout" may be of some assitance.

@@

Another regularization technique addresses the parameters in the encoder and the
decoder. In our models, both of these layers have the vocabulary size as one of
their dimensions. The encoder transforms the (vocabulary sized) one-hot encoding
into a word encoding while the decoder transforms the hidden state into a
softmax as big as the vocabulary size. These sizes can get huge as the
vocabulary can contain from 20k to 800k with languages with many compound words
(such as Finnish), which means that overfitting is quite likely as many of the
parameters included in encoder and decoder may be redundant.

In order to help solving this issue, Press O. et al \cite{press16} proposed
sharing the same weights for encoder and decoder, as one is the same size as the
transpose of the other and both perform a translation between a dense encoded
layer and a sparse one. They showed in their work how this technique can reduce
the number of parameters in the model without hurting the performance while
lowering the perplexity score.

@@

The starting values for the model parameters play a significant role in the
training process for RNNs. The reason is mainly due to the depth these models
reach while propagating through time. As the hidden state travels from time step
to the next, an activation function is applied. The compound effect of
performing the activation function several times tend to take the hidden state
value towards the saturation zone of the activation functions. In order to avoid
the saturation of the hidden units, the starting values must be chose carefully
to ensure a correct behavior from the beginning, starting the training with as
many active units as possible.

@@

Xavier initialization \cite{xavier10} was designed for the sigmoid and tanh
functions. This is the initialization relevant to us since we have applied such
functions more than the RELU. Xavier initialization consists on starting the
hidden state values as a random distribution with variance as follows: $$var =
\sqrt{\frac{1}{size_h}}$$

@@

We measured differences in performance with 60, 90, 150 and 200 hidden units per
RNN cell for each model architecture with 1 RNN cell, trained
to five epochs in batches of 200 sentences at a time.

@@

On both LSTM and GRU models, any use of encoder embeddings noticably outperformed
one-hot encoding on almost all metrics, except METEOR for LSTMs. It seems to be
the case that the learned embeddings encoded useful information about the
non-temporally dependent relationships between words. For Elman RNN cells, we
again observed that because training was so unstable, embeddings did not
appear to make much of a difference to accuracy, even if they result in
much lower computational complexity.

@@

For the implementation of Tied Weights we were expecting lower perplexity levels
as well as a small improvement or no change on the other metrics, however the
LSTM and GRU models didn't show any perplexity improvements while the other
metrics got worse. Our implementation of tied weights had a negative impact. For
the tanh model as small perplexity improvement is shown but the other metrics
are very unstable to draw any conclusions from them.

@@

Experimentation with model hyperparameters that increase or decrease information
retention (such as number of hidden layers or their dimensionality) peformed in line
with expectations. In general, increasing the amount of information stored by the network
allowed it to generate more marginally more realistic language with a lower degree of
perplexity at the output layers.

Usage of trainable word embeddings at the encoding layers outperformed our expectations -
we found that there were no situations where using word embeddings did not result in both
a computational performance and natural language generation performance improvement. In fact,
word embeddings resulted in the most significant perforamnce improvement out of any of the
modifications that we made to the baseline models. We also found that using output shortlists
of length 5000 or above for the WikiText models resulted in scores competitve with the baseline
on both METEOR and ROUGE, but with twice to fource times less computational complexity because
not as many dot products need to be performed at the output layer for each word prediction. However,
if one had to choose between shortlisting and trainable word embeddings at the input layer, embeddings
are able to reduce compute complexity much more whilst also improving accuracy, something that
did not happen with shortlists at the output layer.

@@

@@

\bibliographystyle{plain}
\bibliography{bibliography.bib}
\end{document}

>>>>

Prior to the introduction of RNNs, most language models used n-grams
in Hidden Markov Model chains, where it was assumed that the $k$th
word depended only on the prior $n$ words. One key problem
with this model is that the context window is of a fixed size, leading
to a horizon problem. The horizon problem occurs because if there were a word
in the word sequence which is extremely predictive of the $(n + 1)$th
word away from the current word, that word would simply not be taken
into account in that model. Worse still, the fixed size context window
means that unnecessary weight is put on words that happen to be in the context
window that may actually not be all that predictive were the context window
infinite.

In contrast, RNNs seek to solve this problem by introducing a Neural Network
architecture with recurrent connections, meaning that as the RNN makes predictions
over a sequence observations, it takes into account a hidden layer produced
by the previous observation (which in turn was a product of all the observations
before that). In effect, the RNN architecture encodes the contextual history
into the hidden layer vector allowing for an encoded approximation of infinite
contextual history.

@@

We build on top of the paper by performing experiments with new model types
and improvements and measure their effectiveness against certain well-known datasets.

@@

The first dataset we use is the WikiText dataset. This is a smaller dataset than
Gigaword and WSJ, but allowed us to
iterate faster and run more experiments.

@@

$$ WER = \frac{S + D + I}{N} $$, where $N$ is the number of
words and $S, D, I$ are substitutions, deletions, insertions.
This is similar to the Levenshtein Distance between documents (Haldar et al. \cite{haldar11}).

However, such a metric doesn't take into account the fact that
a recognized sequence may differ in length to a reference
sequence.

Improving on this situation, ROUGE measures based on the
longest matching sub-sequences, overlapping pairs and n-gram
co-occurrences.

\subsection{METEOR}
\label{sec:meteor}
Another metric trying to improve BLEU is
METEOR. The novel thing about this metric is the use of Recall as well as
Precision to compute the score. It computes the harmonic mean of Recall and
Precision over uni-grams, the weight for recall being higher than Precision,
meaning that false negatives (having a low probability for the true word we are  trying
to predict) will have a big impact on the metric (the score will be lower as the
amount of false negatives increase).

Formally, Precision over an uni-gram is computed as: $$P={\frac
{m}{w_{{t}}}}$$ Where m is the number of uni-grams in the candidate translation
that are also found in the reference translation, and $w_t$ is the number of
uni-grams in the candidate translation.

@@

Unfortunately, Elman Network RNNs are not able to retain all the optimal
contextual in practice. The reasons for this are given by Bengio\cite{Bengio94},
but in summary the issue is that as we backpropagate through time the gradients
from states "further back" in the history start to get smaller and smaller, meaning
that they can be easily disrupted by noise or newer information. This means
that the hidden layer doesn't retain long-term information since it is easily
disrupted by new information.

@@

The LSTM cell uses four "gates" to manage the state of the memory. Each of these
gates comprises a set of weights and activations, described below \cite{colah2015}.

@@

The last gate is an "output" gate which decides which parts of the hidden state are
taken on account to compute the output. The hidden state is first activated by sigmoid to determine
the mask ($o_t$), then the hidden layer is activated by tanh to push the values between $[-1, 1]$, then
finally multiplied by the mask $o_t$ in order to determine the output of the network.

The end result with such a variation is that the hidden state is "protected" from perturbation
by virtue of being combined through a single linear layer and activation layer with the input
state on every time step. Instead, the hidden state is more "explicitly" modified when the
optimized input and forget gates explicitly indicate that information should be added or removed
to the state \cite{hochreiter97}.

@@

Note that the hidden state accuracy benefits here come with a computational complexity cost - every
LSTM cell has three internal linear layers and five activations, which we pay for both in the forward
pass and in backpropagation through time.

@@

Traditionally, the Language Models were fed each word as a One-Hot encoded vector as large as the
vocabulary is. Since vocabularies can be really large (20k to 800k for languages
with many compounded words such as Finnish), this presents a computational
problem, matrix multiplication can get very  expensive quite fast as the size of the matrices being multiplied grow. Word Embeddings provide an
alternative procedure to One-Hot in which the words to be fed to the network are encoded
in an $n$ dimensional space. The embedding layer can learn a word encoding such
that words semantically related to each other, get similar encoded vectors after
the transformation. Even some basic arithmetic is possible by adding and subtracting the euclidean distance of vectors. For example, if you
subtract (the vector for "queen" minus the vector from "woman") from the vector for "king" you get the vector
for "man".

@@

Word embeddings also have the distinct advantage that they can reduce both training
and inference complexity because fewer multiplications need to be done at
the RNN, LSTM or GRU cell phase both at the inference and training time. The
computation of the embedding layer can be heavily optimized because the relevant
input will only have a non-zero value in a single vector component.

@@

The overall theory behind shortlists is that any optimal model operates
on a Pareto principle, 20% of the words which are the most frequent will
be predicted roughly 80% of the time. Since the computation of the output
encodings of height $|V|$ is a fixed cost on every training and inference
cycle, the user of the model is paying for the capacity to predict words which
will seldom be predicted. A good shortlist should be able to substantially
reduce the computational complexity of the model whilst still retaining a
relatively high degree of accuracy.

@@

The Milkolov paper flags that performance on the validation could not be
improved by using "standard regularization procedures" (we assume that
the paper must have been referring to L1 or L2 regularization based on the
techniques in existence at the time) \cite{Milkolov10}. We propose that a
popular regularization technique known as "Dropout" may be of some assistance.

@@

Another regularization technique addresses the parameters in the encoder and the
decoder. In our models, both of these layers have the vocabulary size as one of
their dimensions. The encoder transforms the (vocabulary sized) one-hot encoding
into a word embedding while the decoder transforms the hidden state into a
softmax as big as the vocabulary size. These sizes can get huge as the
vocabulary can contain from 20k to 800k, which means that overfitting is quite likely as many of the
parameters included in encoder and decoder may be redundant.

In order to help solving this issue, Press O. et al. \cite{press16} proposed
sharing the same weights for encoder and decoder, as one is the same size as the
transpose of the other and both perform a translation between a dense encoded
layer and a sparse one. They showed in their work how this technique can reduce
the number of parameters in the model without hurting the performance while
lowering the perplexity score.

@@

The starting values for the model parameters play a significant role in the
training process for RNNs. The reason is mainly due to the depth these models
reach while propagating through time. As the hidden state travels from time step
to the next, an activation function is applied. The compound effect of
performing the activation function several times tend to take the hidden state
value towards the saturation zone of the activation functions. In order to avoid
the saturation of the hidden units, the starting values must be chose carefully. The goal is then 
to ensure a correct behavior from the beginning, starting the training process with as
many active units as possible.

@@

Xavier initialization \cite{xavier10} was designed for the sigmoid and tanh
functions. This is the initialization relevant to us since we have applied such
functions more than the RELU alternative. Xavier initialization consists on starting the
hidden state values as a random distribution with variance as follows: $$var =
\sqrt{\frac{1}{size_h}}$$

@@

We measured differences in performance with 60, 90, 150 and 200 hidden units per
RNN cell for each model architecture with 1 RNN layer, trained
to five epochs in batches of 200 sentences at a time.

@@

On both LSTM and GRU models, any use of encoder embeddings noticeably outperformed
one-hot encoding on almost all metrics, except METEOR for LSTMs. It seems to be
the case that the learned embeddings encoded useful information about the
non-temporally dependent relationships between words. For Elman RNN cells, we
again observed that because training was so unstable, embeddings did not
appear to make much of a difference to accuracy, even if they result in
much lower computational complexity.

@@

For the implementation of Tied Weights we were expecting lower perplexity levels
as well as a small improvement or no change on the other metrics, however the
LSTM and GRU models didn't show any perplexity improvements while the other
metrics got worse. Our implementation of tied weights had a negative impact. For
the tanh model a small perplexity improvement is shown but the other metrics
are too unstable to draw any conclusions from them.

@@

Experimentation with model hyperparameters that increase or decrease information
retention (such as number of hidden layers or their dimensionality) performed in line
with expectations. In general, increasing the amount of information stored by the network
allowed it to generate (marginally) more realistic language with a lower degree of
perplexity at the output layers.

Usage of trainable word embeddings at the encoding layers outperformed our expectations. We found that there were no situations where using word embeddings did not result in both
a computational performance and natural language generation performance improvement. In fact,
word embeddings resulted in the most significant performance improvement out of any of the
modifications that we made to the baseline models. We also found that using output shortlists
of length 5000 or above for the WikiText models resulted in scores competitve with the baseline
on both METEOR and ROUGE, but with twice to four times less computational complexity because
not as many dot products need to be performed at the output layer for each word prediction. However,
if one had to choose between shortlisting and trainable word embeddings at the input layer, embeddings
are able to reduce compute complexity much more whilst also improving accuracy, something that
did not happen with shortlists at the output layer.

@@

Additionally, it would be interesting to apply additional model architectures such as Bidirectional RNNs and Attention-based models. In particular, the latter have shown great performance improvements during the last years, achieving for example a new state-of-the-art result in machine translation in 2015 (Luong et al. \cite{luong15}) and also matching state-of-the-art results in Language Modeling with more than 4 times less parameters in comparison with other models (Salton et al. \cite{salton17}).

@@

\bibliographystyle{plain}
\bibliography{bibliography}
\end{document}
